Fusion Moves for Correlation Clustering

### 2014-11-14 paper/egpaper_for_review.tex
fixed stuff

Fusion moves as defined in~\cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning, \cf Fig.~\ref{fig:fusion_abiguty}.
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{E}_0^{\breve{y}}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}_{ij}    & = \max\{ y_{ij}', y_{ij}''\}  & \forall {ij}\in E\\  % y' \cup y'' \\
\mathcal{E}_0^{\breve{y}}  & =  \{ ij \in E \; | \; \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{E}_0^{\breve{y}}$.
%
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
      &\quad \textrm{s.t.} \quad y_{ij} = 0 & \forall (i, j) \in \mathcal{E}_0^{\breve{y}} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy w.r.t. the proposal
$y'$ and $y''$,
because $y'$ and $y''$ are feasible solutions for problem~\ref{eq:fusion_move_a}.
%\begin{align}
%  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
%\end{align}
%

@@

As classical fusion, CC-Fusion does not provide a lower bound on the objective and 
has no sound stopping condition.
For the latter we use a maximal number of iterations and maximal number of iteration without improvement as stopping condition.

@@

\begin{figure}
\centering
\input{polytope-tikz.tex}
\caption{Each fusion move can be interpreted as an optimization of an inner polytope.
Each inner polytope includes the current vertex. Starting with $y_0$ we optimize over the red polytope 
and find $y_1$ as optima. When optimize over the blue polytope we stay in $y_1$ as optima.
When optimizing over the green polytope we find $y_2$ which we will never leave again or any polytope containing~$y_2$. 
}
\label{fig:polyheadral}
\end{figure}

@@

As discussed in~\cite{Lempitsky-2010}, proposals
should have two properties: \emph{quality} 
and \emph{diversity}.

@@

\noindent \textbf{Randomized Hierarchical Clustering (RHC):}
%\subsubsection{Randomized HierarchicalClustering (RHC)}
%
To generate fast energy aware proposals we can use energy
based hierarchical clustering (EHC) as defined in Sec.~\ref{sec:ehc}.
EHC follows the energy function, therefore the \emph{quality}
of the proposals is high.
To get \emph{diversity} among the different proposal,
we add normally distributed noise  $\mathcal{N}(0, \sigma_{ehc})$
to each edge weight.
To get proposals of the desired \emph{size}, we use a
different stop condition as in EHC, and stop only if a certain number of connected
components is reached.

@@

\noindent \textbf{Randomized Watersheds (RWS):}
%\subsubsection{Randomized Watersheds (RWS)}
Watersheds have become quite popular for
graph segmentation and have a strong connection
to energy minimization~\cite{couprie_2011}.
%
The edge weighted watershed algorithm~\cite{meyer_2013}
with random seeds can be used to get
cheap proposals.
To improve \emph{quality} we do not use $n$ seeds distributed uniformly
over all nodes but use the following.
%
We draw $n/2$ negative edges, and assign different seed to the endpoints
of each edge.
Doing so, a random subset of negative edges is forced
to be cut within each proposal.
For additional \emph{diversity}, noise $\mathcal{N} (0, \sigma_{ws})$
is added to the edge weighs \cite{straehle_2012}.

>>>>

Fusion moves as defined in~\cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning, \cf Fig.~\ref{fig:fusion_abiguty}.
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{E}_0^{\breve{y}}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}_{ij}    & = \max\{ y_{ij}', y_{ij}''\}  & \forall {ij}\in E\\  % y' \cup y'' \\
\mathcal{E}_0^{\breve{y}}  & =  \{ ij \in E \; | \; \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{E}_0^{\breve{y}}$.
%
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
      &\quad \textrm{s.t.} \quad y_{ij} = 0 & \forall (i, j) \in \mathcal{E}_0^{\breve{y}} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy w.r.t. the proposals
$y'$ and $y''$,
because $y'$ and $y''$ are feasible solutions for problem~\ref{eq:fusion_move_a}.
%\begin{align}
%  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
%\end{align}
%

@@

As classical fusion, CC-Fusion does not provide a lower bound on the objective and 
has no sound stopping condition.
For the latter we use a maximal number of iterations and maximal number of iterations without improvement.

@@

\begin{figure}
\centering
\input{polytope-tikz.tex}
\caption{Each fusion move can be interpreted as an optimization of an inner polytope.
Each inner polytope includes the current vertex. Starting with $y_0$ we optimize over the red polytope 
and find $y_1$ as optimum. When optimizing over the blue polytope we stay in $y_1$ as optimum,
when optimizing over the green polytope we find $y_2$ which we will never leave again. 
}
\label{fig:polyheadral}
\end{figure}

@@

As discussed in~\cite{Lempitsky-2010}, proposals
should have two properties: \emph{high quality} 
and \emph{large diversity}.

@@

\noindent \textbf{Randomized Hierarchical Clustering (RHC):}
%\subsubsection{Randomized HierarchicalClustering (RHC)}
%
To generate fast energy aware proposals we can use energy
based hierarchical clustering (EHC) as defined in Sec.~\ref{sec:ehc}.
EHC follows the energy function, therefore the \emph{quality}
of the proposals is high.
To get \emph{diversity} among the different proposal,
we add normally distributed noise  $\mathcal{N}(0, \sigma_{ehc})$
to each edge weight.
To get proposals of the desired \emph{size}, we use a
different stop condition for EHC, and stop only if a certain number of connected
components is reached.

@@

\noindent \textbf{Randomized Watersheds (RWS):}
%\subsubsection{Randomized Watersheds (RWS)}
Watersheds have become quite popular for
graph segmentation and have a strong connection
to energy minimization~\cite{couprie_2011}.
%
The edge weighted watershed algorithm~\cite{meyer_2013}
with random seeds can be used to find
cheap proposals.
To improve \emph{quality} we do not use $n$ seeds distributed uniformly
over all nodes but use the following.
%
We draw $n/2$ negative edges, and assign different seeds to the endpoints
of each edge.
Doing so, a random subset of negative edges is forced
to be cut within each proposal.
For additional \emph{diversity}, noise $\mathcal{N} (0, \sigma_{ws})$
is added to the edge weighs \cite{straehle_2012}.
